---
title: "Filtering Noise in the Data"
author: "Dave Hurst"
date: "Thursday, December 25, 2014"
output: html_document
---

Preliminary notes for dealing with noise in the data.

```{r get_driver, echo=FALSE}
source("telematic_util.R")

data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"

drivers.dir <- dir(data.dir)
set.seed(2349)

random.driver <- order(runif(1:length(drivers.dir), ))

N.DRIVERS <- 1
driver.df <- data.frame(id=integer()
                        , speed.avg=numeric(), speed.max=numeric()
                        , break.max=numeric(), accel.max=numeric())
for (i in 1:N.DRIVERS) {
    driver.id <- drivers.dir[random.driver[i]]
    j <- 99  #grab
    trip <- getTrip( driver.id, j )
    driver.df[i, "id"] = driver.id
    driver.df[i, "speed.avg"] =  mean(trip$v)
    driver.df[i, "speed.max"] =  max(trip$v)
    driver.df[i, "break.max"] =  min(trip$a)
    driver.df[i, "accel.max"] =  max(trip$a)
}    
    
```
Now we've got the `r j`th trip for driver `r drivers.dir[random.driver[i]]` in memory. 

Here's an obvious depiction of the problem.  The plot below shows the last 90 seconds of the journey.  It's clear that the car is motionless, and yet we get wiggle in the coordinates that show up as 21 m/s velocity.

```{r show_wiggle}
plotTrip(trip, tmin=nrow(trip)-90, v.mark=1)
```

Plot of cumulative distance (note: trip$v can be used as segment distance, since t=1), and pick an arbitrary region to zoom in on (denoted with a red box)

```{r cum_dist}
plot(cumsum(trip$v)/1000, ylab="km", main= "Cumulative Distance", xlab="seconds", type="l")
t1 <- 790; t2 <- 810
d1 <- cumsum(trip$v)[t1] - 500 
d2 <- cumsum(trip$v)[t2] + 500
box.x <- c(t1, t2, t2, t1, t1)
box.y <- c(d1, d1, d2, d2, d1)/1000
lines(box.x, box.y, col="red")
```   

Zoom in (on the red box above) to see the noise.  This plot is basically saying that car shifted 1 meter in its lane, which is plausible, but could be GPS error.
```{r zoom_dist}
plot(t1:t2, cumsum(trip$v)[t1:t2]/1000, ylab="km", main= "Cumulative Distance", xlab="seconds",type="l")
```

