---
title: "Manual Trip Matching for Driver 2591"
author: "Dave Hurst"
date: "Friday, March 6, 2015"
output: html_document
---

```{r, echo=FALSE}
library(knitr)
source("telematic_util.R")
data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"
segs.dir <- "profile_segs"

driver.id <- 2591

route.unk <- 1:200

route.a <- c( 4, 10, 11, 16, 18, 37, 40, 43, 44, 45, 47, 49, 52, 53, 54, 63, 70, 80, 82, 90, 104, 105
              , 107, 111, 115, 116, 117, 120, 121, 123, 130, 131, 138, 140, 142, 146, 149, 152, 154, 155
              , 157, 162, 178, 180, 184, 189, 192, 198, 200)

route.b <- c( 14, 67, 69, 74, 87, 181, 182, 191, 193)

route.unk <- route.unk[ -c(route.a, route.b)]

```

#Route A

Plotting the group of trips I'll call Route A
```
```{r routea, fig.width=10}
print( route.a )
par.orig <- par(mar=c(1,1,1,1), mfrow=c(1,2), ask=FALSE)
for (trip.id in route.a) {
    plotTrip( getTrip(driver.id, trip.id), header=FALSE)
}
```

Notes:
- 149 has a turn off the others don't

## Route B

```{r routeb, fig.width=10}
print( route.b )
par(mar=c(1,1,1,1), mfrow=c(1,2), ask=FALSE)
for (trip.id in route.b) {     plotTrip( getTrip(driver.id, trip.id), header=FALSE)
}
```

## Everything Else
```{r routeunk, fig.width=10}
print( route.b )
par(mar=c(1,1,1,1), mfrow=c(1,2), ask=FALSE)
for (trip.id in route.unk) {     plotTrip( getTrip(driver.id, trip.id), header=FALSE)
}
par(par.orig)
```


