---
title: "RHadoop Benchmark"
author: "Dave Hurst"
date: "Wednesday, March 18, 2015"
output: html_document
---

```{r, echo=FALSE}
# set R environments
Sys.setenv(HADOOP_CMD="/home/hadoop/bin/hadoop")
Sys.setenv(HADOOP_STREAMING="/home/hadoop/contrib/streaming/hadoop-streaming.jar")
Sys.setenv(JAVA_HOME="/usr/java/latest/jre")

library(rmr2);  
library(jsonlite)

map.driver <- function(k, v) {   
    log <- sprintf("%s: begin map.drive with %d drivers\n", Sys.time() , length(v))
    #     
    #trips.json <- list()
    trip.keys <- character()
    p0 <- p1 <- proc.time()
    done <- integer()
    
    trips <- data.frame()
    for (driver.id in v) {
        log <- paste0( log, sprintf("%d: ", driver.id))
        for (trip.id in 1:N.TRIPS) {
            trip.key <- paste0(driver.id, ".", trip.id)
            trip <- getTrip( driver.id, trip.id, v.thresh=2, data=data)
            trips <- rbind(trips, trip)
            #trip.json <- toJSON( trip )
            trip.keys <- c(trip.keys, trip.key)
            #trips.json <- c(trips.json, trip.json)
        }
        p2 <- proc.time()
        log <- paste0( log, sprintf(",%d", round((p2-p1)[3]) ) )
        p1 <- p2
        
        done <- c(done, driver.id)
    }
    final.time <- round((p1-p0)[3])
    log <- paste0( log
                   , sprintf("\nCompleted drivers (%d secs): ", final.time)
                   , paste0(done, collapse=","), "\n")
    #keyval( c("log", trip.keys), c(log, trips.json) )
    #keyval( c("log", "json_size"), c(log, object.size(trips.json)) )
    keyval("head(trips)", head(trips))
}
getTrip <- function(driver.id, trip.id, v.thresh=5, data=NULL) {
    
    bearing.smooth <<- 0
    bearing.last   <<- 0
    trip <- with(data, data[driver==driver.id & tripnum==trip.id, 1:2] )  

    trip.last <- trip[-nrow(trip), ]
    trip <- trip[-1, ]   #first derivative 
    trip$x.d <- trip$x - trip.last$x
    trip$y.d <- trip$y - trip.last$y
    trip$v <- sqrt( trip$x.d^2 + trip$y.d^2 )  # distance travelled per second
    trip.last <- trip[-nrow(trip), 3:5 ]
    trip <- trip[-1, ]    #2nd derivative 
    trip$x.d2 <- trip$x.d - trip.last$x.d
    trip$y.d2 <- trip$y.d - trip.last$y.d
    trip$a <- trip$v - trip.last$v     
    bearing <- numeric()
    bearing.smooth <<- 0
    bearing.last   <<- 0
    for(i in 1:nrow(trip)) {
        bearing[i] <- ifelse( trip[i,]$v > v.thresh, calcBearing( trip[i,]), NA )
    }
    trip <- cbind(trip, bearing)    
    
    trip$driver <- driver.id
    trip$tripnum <- trip.id
    
    return(trip)
}
calcBearing <- function( t, smooth=TRUE ) {
    #smooth causes a 2deg ccw rotation at 179 to go to 181 rather than -179
    #, but results in angles > 360  ... may be imperfect if there are large swings in bearings
    vx <- t$x.d
    vy <- t$y.d
    bearing <- atan(vy / vx) * 180/pi
    if ( vx  < 0 )  {
        if ( vy < 0 ) {   #3rd Quadrant
            bearing <- bearing - 180  
            if (bearing.last %% 360 <= 180) bearing.smooth <<- bearing.smooth + 360
        } else {          #2nd Quadrant
            bearing <- bearing + 180
            if (bearing.last %% 360 > 180) bearing.smooth <<- bearing.smooth - 360
        }
    }
    if (smooth) bearing <- bearing + bearing.smooth
    if ( ! is.nan(bearing) ) bearing.last <<- bearing  
    return(bearing)
}
```

This benchmark compares feature extraction time for a subset of the kaggle telematics data in 3 separate run environments.  The benchmark is run in an AWS EMR cluster and uses 3 configurations:
- R in native mode running on the master node
- R with rmr running in local mode on the master node
- R with rmr running in hadoop mode on the full cluster

File Dependencies:
- driver_2591_data.RData

Load the data and print a few head and tail lines
```{r load_data}
load(file= "driver_2591_data.RData")
head(data, 4)
tail(data, 4)

drivers = rep(2591, 5)  # process the same driver 5 times (approx 5 different drivers)
N.TRIPS <- 200  # 200 for full set of trips, but reduce for dev/debug
```

## Test1 - R Native Environment
```{r test1}
p1 <- proc.time()
trips <- data.frame()
for (driver.id in drivers) {
    for (trip.id in 1:N.TRIPS) { 
        trips <- rbind(trips, getTrip( 2591, trip.id , v.thresh=2, data=data)) 
    }
}
p2 <- proc.time()
benchmark.times <- (p2-p1)[3]

head(trips)
object.size(trips)
benchmark.times
rm(trips)
```

## Test2 - rmr in local mode
```{r test2}
rmr.options(backend="local")

p1 <- proc.time()
out <- mapreduce(
    input = to.dfs( drivers ),
    map = map.driver 
    )
p2 <- proc.time()
benchmark.times <- c( benchmark.times, (p2-p1)[3] )
values(from.dfs(out))
benchmark.times
```
## Test3 - rmr in hadoop mode
```{r test3}
rmr.options(backend="hadoop")

p1 <- proc.time()
out <- mapreduce(
    input = to.dfs( drivers ),
    map = map.driver 
    )
p2 <- proc.time()
benchmark.times <- c( benchmark.times, (p2-p1)[3] )

values(from.dfs(out))
benchmark.times

names(benchmark.times) <- c("Native R", "rmr LOCAL", "rmr HADOOP")
barplot(benchmark.times, ylab="Time (sec)")
```

## Appendix - functions used in this benchmark
```{r apx}
map.driver
getTrip
calcBearing
```