---
title: "Segmentation of Trip Data - Part4"
author: "Dave Hurst"
date: "Tuesday, December 30, 2014"
output: pdf_document
---
```{r init, echo=FALSE}
source("telematic_util.R")

data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"

drivers.dir <- dir(data.dir)
set.seed(2349)

random.driver <- order(runif(1:length(drivers.dir), ))
driver.df <- data.frame(id=integer()
                        , speed.avg=numeric(), speed.max=numeric()
                        , break.max=numeric(), accel.max=numeric())

```
## Build trip profiles

I wrote `tripProfile` to count the number of *straight, accelerating*, and *decelerating* segments along with some characteristics (see at the bottom).

Here's what it comes up with for the driver we've been working with:
```{r prof_summary}
driver.id <- 2591

trip.prof <- data.frame()
ptime0 <- proc.time()
for(i.trip in 1:200) {
    trip <- getTrip( driver.id, i.trip, v.thresh=1 )    #v.thresh = lowered for slow trips
    tp <- getTripProfile( trip )
    trip.prof <- rbind(trip.prof, tp)
    cat(i.trip, ": length, dist, straights, accels, decels # ", tp$trip.len, round(tp$trip.dist/1000,1), tp$ss.n, tp$acc.n, tp$dec.n, "...\n")
}
rownames(trip.prof) <- 1:200
ptime1 <- proc.time()
print (ptime1 - ptime0)

```

## Notes / Observations

### Trip 7
Trip 7 shows a break in the deceleration (downshift?).  This can be smoothed out by increasing the MA value (I added a parameter for `segment.parse.XXcel`), but do we want to?  Notice the value for *a.mid* (the midpoint deceleration value changes significantly).

One idea (for parse.decel) would be to pre-process the segment with a very high MA to get the boundaries of the segment, but process it with better resolution.  It's interesting for this graph that both segments are steeper than the smoothed segment, but the fact that there is a "stutter" could be predictive, so we probably don't want lose that (but will for now).
```{r trip7}
i.trip=7
trip <- getTrip( driver.id, i.trip, v.thresh=1 )  
#tp <- getTripProfile( trip )  #this originally had a warning message which lead me to look at this trip
par.orig <- par(mfrow=c(1,2))
plotTripSegment.speed(trip,1,999)
plotTripSegment.speed(trip,1,999, ma=13)
par(par.orig)
segment.parse.decel(trip) #default ma=5
segment.parse.decel(trip, ma=13)


```

### Trip 72
```{r trip72}
#with(trip.prof, print(trip.prof[ss.n * acc.n * dec.n == 0,  c("ss.n", "acc.n", "dec.n")]))
miss.1   <- with(trip.prof, print(which(ss.n * acc.n * dec.n == 0 )))
miss.all <- with(trip.prof, print(which(ss.n + acc.n + dec.n == 0 )))
cat ("Trips missing at least one segment type:\n", miss.1, "\n")
cat ("Trips with no segment type:\n"             , miss.all, "\n")

i.trip <- 72
tp <- trip.prof[i.trip, ]
cat(i.trip, ": length, dist, straights, accels, decels # ", tp$trip.len, round(tp$trip.dist/1000,1), tp$ss.n, tp$acc.n, tp$dec.n, "...\n")
trip <- getTrip( driver.id, i.trip, v.thresh=1 )  
plotTrip(trip)
```

This looks like a parked car, and is probably all noise.  We should filter trips like these out.
```{r}
trip.prof <- trip.prof[-miss.all, ]
print(summary( trip.prof ))

```
```{r}
trip.id <- rownames(trip.prof[which.max(trip.prof$trip.len), ])
```

## Longest Trip = `r trip.id`
Just for grins, let's take a look at the trip with the max length
```{r}
tp <- trip.prof[trip.id, ]
cat(trip.id, ": length, dist, straights, accels, decels # ", tp$trip.len, round(tp$trip.dist/1000,1), tp$ss.n, tp$acc.n, tp$dec.n, "...\n")
trip <- getTrip( driver.id, trip.id, v.thresh=1 )  
plotTrip(trip)
plotTripSegment6(trip,1,9999)
```

```{r}
nfeat <- with(trip.prof, ss.n + acc.n + dec.n  )
trip.id <- rownames(trip.prof[which.max(nfeat), ])
```

; and again for 
## Trip with most features = `r trip.id`
Just for grins, let's take a look at the trip with the max length
```{r}
tp <- trip.prof[trip.id, ]
cat(trip.id, ": length, dist, straights, accels, decels # ", tp$trip.len, round(tp$trip.dist/1000,1), tp$ss.n, tp$acc.n, tp$dec.n, "...\n")
trip <- getTrip( driver.id, trip.id, v.thresh=1 )  
plotTrip(trip)
plotTripSegment6(trip,1,9999)
```

```{r writeDriverProf}
prof.fn <- sprintf("driver_%d_prof.RData", driver.id);
save(file=prof.fn, trip.prof)
```



```{r prof_code}
getTripProfile
```
