---
title: "Segmentation of Trip Data - Part3"
author: "Dave Hurst"
date: "Thursday, December 25, 2014"
output: html_document
---
```{r init, echo=FALSE}
source("telematic_util.R")

data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"

drivers.dir <- dir(data.dir)
set.seed(2349)

random.driver <- order(runif(1:length(drivers.dir), ))
driver.df <- data.frame(id=integer()
                        , speed.avg=numeric(), speed.max=numeric()
                        , break.max=numeric(), accel.max=numeric())

```
## Investigate Trips with no segments

### A note on the speed plot:

See Part2 for background.

Note ... this is starting to get slow, and I can't get caching to work.

```{r seg0}

driver.id <- 2591
seg.summary <- data.frame( n.seg=integer(), trip.len=integer(), speed.avg=numeric(), speed.max=numeric(),
                           seg.min=numeric(), seg.max=numeric(), seg.avg=numeric(), seg.sd=numeric())
for(i.trip in 1:200) {
    #cat(i.trip,".")
    trip <- getTrip( driver.id, i.trip, v.thresh=5 )    #v.thresh = default
    ss <- segment.parse.bearing(trip)
    seg.summary[i.trip, ]$trip.len <- nrow(trip)
    seg.summary[i.trip, ]$speed.avg <- mean(trip$v)
    seg.summary[i.trip, ]$speed.max <- max(trip$v)
    seg.summary[i.trip, ]$n.seg   <- nrow(ss)
    if ( nrow(ss) > 0) {
       seg.summary[i.trip, ]$seg.min <- min(ss$tlen)
       seg.summary[i.trip, ]$seg.max <- max(ss$tlen)
       seg.summary[i.trip, ]$seg.avg <- mean(ss$tlen)
       seg.summary[i.trip, ]$seg.sd  <- sd(ss$tlen)
    }
}
print(summary(seg.summary))

seg0 <- which(seg.summary$n.seg == 0)
seg0
```

### Trips with 0 segs

Here's an example of the plots for these:
```{r}
trip.id <- 16
trip <- getTrip( driver.id, trip.id )
plotTrip(trip)
plotTripSegment(trip)
plotTripSegment6(trip)
```

* Trip 6: Extremely short trip (194 secs),  doesn't meet 30 sec criteria
* Trip 16: 'straight' segment is 29 secs (1 under the threshold)
* Trip 40: Too short 181 secs
* Trip 45... : yada yada yada


This is starting to get redundant, so let's look at all the trips in sorted order (lowest trip time) in order to see what these contain that we could use.

```{r}
asc <- order(seg.summary$trip.len)
print(seg.summary[asc, 1:7])
```

### trip 157 (2.5 mins)
```{r}
trip.id <- 157
trip <- getTrip( driver.id, trip.id, v.thresh=1 )  # added v.thresh to allow for lower speed heading capture
plotTrip(trip, t.mark=10)
plotTripSegment6(trip, 1, 80, t.mark=10)
```

Looking at this trip, and several like it, I've determined we need to just capture acceleration or braking segments, since almost all trips will have those.  To start, I just look for deceleration and acceleration sequence by looking for sequential decreasing or increasing speeds respectively, where the speed differential is > 5 m/s  (NOTE: I used these to improve the red/green acceleration coloring on the trip plots).

Let's stop here, and start moving on to trip profiles in Part4



