---
title: "Segmentation of Trip Data - Part3"
author: "Dave Hurst"
date: "Thursday, December 25, 2014"
output: html_document
---
```{r init, echo=FALSE}
source("telematic_util.R")

data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"

drivers.dir <- dir(data.dir)
set.seed(2349)

random.driver <- order(runif(1:length(drivers.dir), ))
driver.df <- data.frame(id=integer()
                        , speed.avg=numeric(), speed.max=numeric()
                        , break.max=numeric(), accel.max=numeric())

```
## Investigate Trips with no segments

### A note on the speed plot:

See Part2 for background.

Note ... this is starting to get slow, and I can't get caching to work.

```{r seg0}

driver.id <- 2591
seg.summary <- data.frame( n.seg=integer(), trip.len=integer(), speed.avg=numeric(), speed.max=numeric(),
                           seg.min=numeric(), seg.max=numeric(), seg.avg=numeric(), seg.sd=numeric())
for(i.trip in 1:200) {
    cat(i.trip,".")
    trip <- getTrip( driver.id, i.trip )    
    ss <- segment.parse.bearing(trip)
    seg.summary[i.trip, ]$trip.len <- nrow(trip)
    seg.summary[i.trip, ]$speed.avg <- mean(trip$v)
    seg.summary[i.trip, ]$speed.max <- max(trip$v)
    seg.summary[i.trip, ]$n.seg   <- nrow(ss)
    if ( nrow(ss) > 0) {
       seg.summary[i.trip, ]$seg.min <- min(ss$tlen)
       seg.summary[i.trip, ]$seg.max <- max(ss$tlen)
       seg.summary[i.trip, ]$seg.avg <- mean(ss$tlen)
       seg.summary[i.trip, ]$seg.sd  <- sd(ss$tlen)
    }
}
print(summary(seg.summary))

seg0 <- which(seg.summary$n.seg == 0)
seg0
```

### Trips with 0 segs

Here's an example of the plots for these:
```{r}
trip.id <- 16
trip <- getTrip( driver.id, trip.id )
plotTrip(trip)
plotTripSegment(trip, 1, 20000, t.mark=20)
```

* Trip 6: Extremely short trip (194 secs),  doesn't meet 30 sec criteria
* Trip 16: Mirror image of Trip200.  Doesn't meet spead threshhold
* Trip 40: Too short 181 secs
* Trip 45... : yada yada yad


This is starting to get redundant, so let's look at all the trips in sorted order (lowest trip time) in order to see what these contain that we could use.

```{r}
asc <- order(seg.summary$trip.len)
print(seg.summary[asc, ])
```

```{r}
trip.id <- 157
trip <- getTrip( driver.id, trip.id )
plotTrip(trip, t.mark=10)
plotTripSegment(trip, 1, 10000, t.mark=20)
```
