---
title: "Selected Driver Observations"
author: "Dave Hurst"
date: "Sunday, December 21, 2014"
output: pdf_document
---

Playing around with a few driver trip profiles as part of the EDA.  *Caveat:* I haven't read any of the forums at this point.


Here's the data read statement. It's a little overkill, but I'm setting things up for multiple trips/drivers
```{r}
data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"

drivers.dir <- dir(data.dir)
set.seed(2349)

random.driver <- order(runif(1:length(drivers.dir), ))

N.DRIVERS <- 1
driver.df <- data.frame(id=integer()
                        , speed.avg=numeric(), speed.max=numeric()
                        , break.max=numeric(), accel.max=numeric())

```

Here's the code for a single trip.  I'm calculating *speed* and *acceleration* since these will be used for feature creation
```{r}
i <- 1
j <- 1

    driver.dir <- paste(c(data.dir, drivers.dir[random.driver[1]]), collapse="/")
    driver.files <- dir(driver.dir)

        trip <- read.csv( paste(c(driver.dir, driver.files[j] ), collapse="/"))
        trip.last <- trip[-nrow(trip), ]
        trip <- trip[-1, ]
        trip$x.d <- trip$x - trip.last$x
        trip$y.d <- trip$y - trip.last$y
        trip$v <- sqrt( trip$x.d^2 + trip$y.d^2 )  # distance travelled per second
        trip.last <- trip[-nrow(trip), 3:5 ]
        trip <- trip[-1, ]
        trip$x.d2 <- trip$x.d - trip.last$x.d
        trip$y.d2 <- trip$y.d - trip.last$y.d
        trip$a <- trip$v - trip.last$v      

    driver.df[i,"id"] = drivers.dir[random.driver[1]]
    driver.df[i, "speed.avg"] =  mean(trip$v)
    driver.df[i, "speed.max"] =  max(trip$v)
    driver.df[i, "break.max"] =  min(trip$a)
    driver.df[i, "accel.max"] =  max(trip$a)

driver.df[i,]

```

Here's the speed and acceleration profile for the first trip.  *Note:* The R-markdown plots are different than what I'm getting out of R.  I see a spike in the Markdown that isn't in R-Studio, and its making the acceleration plot useless.   The Velocity is also much better if you plot directly from R -- didn't feel like fighting this.

```{r}    
plot(trip$v, type="l", main="Velocity")
plot(trip$a, type="l", main="Acceleration")
```
A couple of things to note in this plot:  For the first half of the trip, the driver seems to be moving on surface streets at around 25 m/s (~55 mph) with several slow downs, but no stops (timed traffic lights? Country roads with no cops?).  the acceleration (slope of the velocity curve) back to the speed limit is consistent, and this probably a characteristic of a particular driver.

Acceleration/Deceleration during "constant" stretches is probably also a characteristic we can capture

Here's the route ... not sure what this tells us.
```{r}
plot(trip[, 1:2], type="l", asp=1, main= "Plot of Route")

```

Same plots and calculation for a different trip
used for feature creation
```{r}
i <- 1
j <- 103
```
```{r, echo=FALSE}

    driver.dir <- paste(c(data.dir, drivers.dir[random.driver[1]]), collapse="/")
    driver.files <- dir(driver.dir)
         trip <- read.csv( paste(c(driver.dir, driver.files[j] ), collapse="/"))
        trip.last <- trip[-nrow(trip), ]
        trip <- trip[-1, ]
        trip$x.d <- trip$x - trip.last$x
        trip$y.d <- trip$y - trip.last$y
        trip$v <- sqrt( trip$x.d^2 + trip$y.d^2 )  # distance travelled per second
        trip.last <- trip[-nrow(trip), 3:5 ]
        trip <- trip[-1, ]
        trip$x.d2 <- trip$x.d - trip.last$x.d
        trip$y.d2 <- trip$y.d - trip.last$y.d
        trip$a <- trip$v - trip.last$v      

    driver.df[i,"id"] = drivers.dir[random.driver[1]]
    driver.df[i, "speed.avg"] =  mean(trip$v)
    driver.df[i, "speed.max"] =  max(trip$v)
    driver.df[i, "break.max"] =  min(trip$a)
    driver.df[i, "accel.max"] =  max(trip$a)

driver.df[i,]

plot(trip$v, type="l", main="Velocity")
plot(trip$a, type="l", main="Acceleration")
plot(trip[, 1:2], type="l", asp=1, main= "Plot of Route")

```

And again for a different driver (a UPS truck?)

```{r}
i <- 62
j <- 72
```
```{r, echo=FALSE}

    driver.dir <- paste(c(data.dir, drivers.dir[random.driver[1]]), collapse="/")
    driver.files <- dir(driver.dir)

        trip <- read.csv( paste(c(driver.dir, driver.files[j] ), collapse="/"))
        trip.last <- trip[-nrow(trip), ]
        trip <- trip[-1, ]
        trip$x.d <- trip$x - trip.last$x
        trip$y.d <- trip$y - trip.last$y
        trip$v <- sqrt( trip$x.d^2 + trip$y.d^2 )  # distance travelled per second
        trip.last <- trip[-nrow(trip), 3:5 ]
        trip <- trip[-1, ]
        trip$x.d2 <- trip$x.d - trip.last$x.d
        trip$y.d2 <- trip$y.d - trip.last$y.d
        trip$a <- trip$v - trip.last$v      

    driver.df[i,"id"] = drivers.dir[random.driver[1]]
    driver.df[i, "speed.avg"] =  mean(trip$v)
    driver.df[i, "speed.max"] =  max(trip$v)
    driver.df[i, "break.max"] =  min(trip$a)
    driver.df[i, "accel.max"] =  max(trip$a)

driver.df[i,]

plot(trip$v, type="l", main="Velocity")
plot(trip$a, type="l", main="Acceleration")
plot(trip[, 1:2], type="l", asp=1, main= "Plot of Route")

```