---
title: "Trip Acceleration Segments Part 1"
author: "Dave Hurst"
date: "Friday, March 6, 2015"
output: html_document
---

```{r, echo=FALSE}
library(knitr)
source("telematic_util.R")
library(car)  #used for boxplots
data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"
```

Working with just acceleration segments, let's see what we can figure out.

Using  the 'getAccels' function, and 'gen_profiles_by_driver.R', I've got acceleration segment data for several drivers.  For now, we'll load 6 of them, which includes our buddy 2591, the driver we'll start trying to isolate.


```{r loadacc, fig.width=10, cache=TRUE}
set.seed(2349)
drivers.dir <- dir(data.dir)
random.driver <- order(runif(1:length(drivers.dir), ))
driver.list <- drivers.dir[random.driver]  #needless complication that I'm stuck with at the moment
accels.dir <- "profile_accels"
segs.dir <-  "profile_segs"

driver.id <- driver.list[1]
accels.fn <- get.fname (accels.dir, driver.id, "accels")
load( file=accels.fn )
segs.fn <- get.fname (segs.dir, driver.id, "segs")
load( file=segs.fn )
asegs <- accels  # this is the data frame we'll get all the acceleration data in
gsegs <- trip.seg # this is the data frame we'll get all the geometry data in

for (driver.id in driver.list[2:6]) {
    accels.fn <- get.fname (accels.dir, driver.id, "accels")
    load( file=accels.fn )
    asegs <- rbind(asegs, accels)
    rm(accels)
    
    segs.fn <- get.fname (segs.dir, driver.id, "segs")
    load( file=segs.fn )
    gsegs <- rbind(gsegs, trip.seg)
    rm(trip.seg)

}
gsegs <- gsegs[ !grepl( "^x.", gsegs$type ), ]  #remove the dead info

head(asegs)
head(gsegs)

```
We currently have data loaded for `r length( unique(asegs$driver))` drivers

Q: How many acceleration segments do we have by velocity threshold  
A: (a.mid has 100% coverage)
```{r nsegs, echo=FALSE}
colSums(! is.na(asegs[,5:9]))

cat("by percentage:\n")
round( 100*colSums(! is.na(asegs[,6:9]))/sum(! is.na(asegs[,5])), 1)
```

Plotting the first segment, we see a curve onto a relatively long straight.
Here are plots through the straight way, and just beyond the accleration.  (note the acceleration phase continues, except for a brief deceleration (gear shift?)).

```{r aseg1, fig.width=10}
trip <- getTrip( 2591, 1)
plotTrip( trip , tmin=21, tmax=125, t.mark=10)
plotTripSegment( trip , tmin=21, tmax=65, t.mark=5, v.mark=2)
```

Q: How many acceleration segments are limited to just 1 or 2 geographical segments and with a straight segment that goes beyond the end of the acceleration by at least 5 seconds?

A: 
```{r segsub}
xx <- with(asegs, asegs[ n.snip <= 2 & strt.overrun >= 5, ])
nrow(xx)  

#what's the number by driver?
table(xx$driver)
```
This looks like enough to do things with by itself.  Let's repeat the velocity threshold question for this subset.
Q: How many acceleration segments do we have by velocity threshold  
A: 
```{r nsegs.xx, echo=FALSE}
colSums(! is.na(xx[,5:9]))

```

What's the next segment look like (From trip 3)

```{r aseg2, fig.width=10}
segx <- xx[2,]
trip <- getTrip( segx$driver, segx$trip)
plotTrip( trip , tmin=segx$snip.t0, tmax=segx$snip.tn, t.mark=10)
plotTripSegment( trip , tmin=segx$t0-5, tmax=segx$tn+10, t.mark=5, v.mark=2)
```

and the next
```{r aseg3, fig.width=10}
segx <- xx[3,]
trip <- getTrip( segx$driver, segx$trip)
plotTrip( trip , tmin=segx$snip.t0, tmax=segx$snip.tn, t.mark=10)
plotTripSegment( trip , tmin=segx$t0-5, tmax=segx$tn+10, t.mark=5, v.mark=2)
```

I think these all look similar enough to move on.  Let's do some EDA on the different features 

```{r box1, fig.width=10, fig.height=10}
par.orig <- par(mfrow=c(5,1), mar=c(2,3,1,1))
for (i in 5:9) { Boxplot( xx[,i] ~ xx$driver, main=colnames(xx)[i]) }
par(par.orig)
```
These don't look terribly differentiated, but there is a weak pattern for the drivers.  Let's try and squeeze a little more juice from the rock:

## Max acceleration for each segment xx)
```{r maxa, fig.width=10}
xx$grp <- "xx"
asegs$grp <- "all"
xx.all <- rbind(asegs, xx)
xx.all$maxa <- apply( xx.all[ ,5:9], 1, max, na.rm=TRUE)
xx.all$mina <- apply( xx.all[ ,5:9], 1, min, na.rm=TRUE)
Boxplot( xx.all$maxa ~ xx.all$driver, main="max a in segment")
Boxplot( xx.all$mina ~ xx.all$driver, main="min a in segment")
```

