---
title: "Trip Sequencing Part 2"
author: "Dave Hurst"
date: "Sunday, February 8, 2015"
output: pdf_document
---

Continuing the work to convert a trip into a sequence of segments.  The input will be a full trip, and the output will be broken up into several components:
- A list of segments with details (or reference to details) on each segment
- A sequence of segment ID's that comprise the trip
- A of set data for each segment type

I'm going to try to reconstruct each trip out of segments made up of the following types:  

* stops *DONE: See part 1* (no driver profile for these, but we need them to describe the segment)  
* sharp turns  
* curves (differentiated from sharp turns by radii (as a function of speed))  
* bends ( differentiated from curves by time < 3 data points)  
* straights  
* uncategorized:  
* - short straights (too short to collect profile info)  
* - slow rolling sections (drive thrus, driveways, etc. -- no profile info)  
* - windy sections (ideally break these up later)  
* - other?  

```{r init, echo=FALSE}
library(knitr)
source("telematic_util.R")

data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"

drivers.dir <- dir(data.dir)
set.seed(2349)
```
Again starting with the familiar 2591/49 trip:
```{r plot6}
driver.id <- 2591
trip.id <- 49
trip <- getTrip( driver.id, trip.id )

plotTripSegment(trip, 1, 99999)
```  

## Sharp Turns

Rules for determining a sharp turn
* R < 20 m
* Direction of curvature must be identical throughout curve ( determined by cross product)

*NOTE: Tunable parameters -* Radius as a function of velocity?

First we need to parse the stops (from Part 1), and break the trip into segments

```{r find.stops }

trip.seg <- segment.by.stops(trip)
trip.seg
```

I created a function to parse turns call segment.parse.turns.  We'll pass each unparsed segments (between stops above) to that to find all the sharp turns.



```{r find.turns }
trip.seg <- segment.by.turns( trip, trip.seg) 
trip.seg
```

The modified `trip.seg` data frame is shown above.  For a little graphical verification, we'll strip out the "turns" and "stops" and overlay those boundaries agains the plot.  It's a little easier to see if we just take a small portion

```{r plot50}
plotTrip(trip, tmax=50, t.mark=5)
segs.1 <- trip.seg[grepl("stop|turn", trip.seg$type) & trip.seg$t0 <50,]
overlaySegmentBorders( trip, c( segs.1$t0, segs.1$tn), size= -1)
segs.1
```

## Functions created/used
```{r turn.func}
segment.clean.points
segment.by.stops
segment.parse.turns
segment.by.turns
```
