---
title: "Trip Sequencing Part 1"
author: "Dave Hurst"
date: "Sunday, January 25, 2015"
output: pdf_document
---

Here I am working to convert a trip into a sequence of segments.  The input will be a full trip, and the output will be broken up into several components:
- A list of segments, with details (or reference to details) on each segment
- A sequence of segment ID's that comprise the trip
- A set data for each segment type

I'm going to try to reconstruct each trip out of segments made up of the following types:  

* stops (no driver profile for these, but we need them to describe the segment)  
* sharp turns  
* curves (differentiated from sharp turns by radii (as a function of speed))  
* bends ( differentiated from curves by time < 3 data points)  
* straights  
* uncategorized:  
* - short straights (too short to collect profile info)  
* - slow rolling sections (drive thrus, driveways, etc. -- no profile info)  
* - windy sections (ideally break these up later)  
* - other?  

```{r init, echo=FALSE}
library(knitr)
source("telematic_util.R")

data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"

drivers.dir <- dir(data.dir)
set.seed(2349)
```
Again starting with the familiar 2591/49 trip:
```{r plot6}
driver.id <- 2591
trip.id <- 49
trip <- getTrip( driver.id, trip.id )

plotTripSegment(trip, 1, 99999)
```  

## Stops and Rolls

Looking at the speed profile there look to be 3 stops during this trip.   Since a stop sign or a drive thru could be a series of stops/slow rolls, and probably won't impart anything in terms of driving style, let's begin the stop segment anytime the speed drops below 1 m/s and continue it until we increases above a roll theshold (which I've arbitrarily set to 2 m/s after some emperical observations)

*NOTE: Tunable parameters -* Stop and Roll threshold speeds 

I wrote a function to find all the stops in a trip based on that criteria.  Here's an example using the current trip followed by the code for the function.

```{r find.stops , results='asis'}
stops <- segment.parse.stops(trip, thresh.stop=1, thresh.roll=2)
stop.id <- 1:nrow(stops)
kable( cbind( stop.id, stops), format = "markdown")
```

(t0 is the data point for the begin of the stop segment, and tn is the end)
```{r}
segment.parse.stops
```

When investigating these points, its useful to see use the PlotTripSegment function.  Just remember to set MA=1 to negate the smoothing, since I'm not applying smoothing for the segmentation analysis.  Here's and example
```{r ex_stop4}
stop <- stops[3,]
plotTripSegment(trip, stop$t0-5, stop$tn+5, ma=1)
```

I add 5 seconds to either side of the segment to see what's going on with it.  In this case, notice we head right back into a stop, which implies the threshold for ending the stop may be too low.  I'll experiment with that below.


Here's an overlay of the trip plot with the location of each segment
```{r plot_stops}
plotTrip.stopOverlay <- function (d,t, ths=1, thr=2) {
    trip <- getTrip(d,t)
    stops <- segment.parse.stops(trip, thresh.stop=ths, thresh.roll=thr)
    print(stops)
    plotTrip(trip)
    for (i in 1:nrow(stops)) {
        t0 <- stops[i, "t0"]; tn <- stops[i, "tn"]
        loc.x <- mean( trip[c(t0,tn), "x"] )
        loc.y <- mean( trip[c(t0,tn), "y"] )
        points( loc.x, loc.y, pch = 5, cex=2)
        text( loc.x, loc.y, labels=i)
    }
}

plotTrip.stopOverlay(driver.id, trip.id)
```

Plot it again with a slightly higher roll threshold, and it looks  cleaner.
```{r}
plotTrip.stopOverlay(driver.id, trip.id, ths=1, thr=2.5)

```

And another for trip 150

```{r}
plotTrip.stopOverlay(driver.id, 150, ths=1, thr=2)
```
Note here that stops 2 and 3 are right on top of each other in what appears to be a Uturn in the segment. So a little investigation is in order.

```{r}
trip.id <- 150
trip <- getTrip(driver.id, trip.id)
stops <- segment.parse.stops(trip, thresh.stop=1, thresh.roll=2)
print(stops)
plotTripSegment(trip, 754, 834 , ma=1, t.mark=10)
```

This shows a stop then a slow roll between 3 and 4 m/s, but that threshold is too high I think, so we'll just have to live with the small segment in between (its about 25 m long).


a good stopping point.



