---
title: "Trip Matching Part 2"
author: "Dave Hurst"
date: "Wednesday, March 25, 2015"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(dplyr);  filter <- stats::filter
source("telematic_util.R")
library(car)  #used for boxplots
data.dir <- "c:/NO_BACKUP/kaggle/telematics/drivers"
trips.dir <- "c:/NO_BACKUP/kaggle/telematics/profile_trips"

quickPlot <- function( driver.id, trip.id, tx=NULL, tl=10, t.mark=20) {
    trip <- getTrip(driver.id, trip.id, v.thresh=2)
    if (is.null(tx)) { tmin=1; tmax=nrow(trip)
    } else { tmin= tx-tl; tmax= tx+tl; t.mark= round(tl/2)}
    plotTrip(trip, t.mark=t.mark, tmin=tmin, tmax=tmax, header=FALSE, v.mark=2)
    segs <- getSegments(trip)
    plotTripOverlay(trip, segs[segs$seg.type=="straight", ], col='yellow') #show straights
    plotTripOverlay(trip, segs[segs$seg.type=="turn", ], col='magenta') #show turns
    plotTrip.stopOverlay(trip, segs[segs$seg.type=="stop", ])
}

```
Although this is labeled "Part 2", there are really two Part 1.  See tripmatching1, and ManualMatch2591.  This is more part 2 to the latter, since I'm going to try to programatically match driver 2591's trips.

start by loading the data into memory
```{r d2591}
driver.id <- 2591
fname <- get.fname( trips.dir, driver.id, "trips")
load(file=fname)
head( with(trip.data,  table(tripnum, seg.type )))

fname <- get.fname( trips.dir, driver.id, "segs")
load(file=fname)
seg.jumps <- seg.all[ seg.all$type == "jump", -1]
seg.stops <- seg.all[ seg.all$type == "stop", -1]
seg.straights <- seg.all[ seg.all$type == "straight", -1]
seg.turns <- seg.all[ seg.all$type == "turn", -1]
rm(seg.all)

```

Let's start with Trip 1. Shown completely on the left, and zoomed in on the tail to the right:

```{r getsegs1, fig.width=10}
par(mfrow=c(1,2))
trips <- trip.data 
trip.id <- 1
trip <- getTrip(driver.id, trip.id, v.thresh=2)
plotTrip(trip)

segs <- getSegments(trip)
plotTripOverlay(trip, segs[segs$seg.type=="straight", ], col='yellow') #show straights
plotTripOverlay(trip, segs[segs$seg.type=="turn", ], col='magenta') #show turns
plotTrip.stopOverlay(trip, segs[segs$seg.type=="stop", ])

plotTrip(trip, tmin=535)
plotTripOverlay(trip, segs[segs$seg.type=="straight", ], col='yellow') #show straights
plotTripOverlay(trip, segs[segs$seg.type=="turn", ], col='magenta') #show turns
plotTrip.stopOverlay(trip, segs[segs$seg.type=="stop", ])

segs
```

- The straights (yellow) look pretty good for this trip (after adjustment -- I set defaults to zone=2, t.thresh=5).  
- There are 3 "turns" (magenta) near the tail end of the trip (a little hard to see without plotting the segments separately).
-- 30<t<40 looks like it should be a turn, but the scale of this trip (3km) is deceiving.  The radius is still quite large, so it qualifies as what I'd classified a "curve", which I'm not breaking out.

The next step is to use some NLP techniques, treating each segment as a "word" in a dictionary.  We'll ignore segments under 10m (stop words), and create n-grams for the trip

Each "word" segment is characterized by its arc length `s.cum`, chord length `s.crow` (shown but not used, since it doesn't differentiate much from `s.cum` -- in theory, the ratio of these should have differentiate between curvey versus straight segments, with the latter being closer to 1:1), angle `d.phi` relative to the preceding segment (`d.phi` is not defined for the first segment), and `seg.type`.  Also capture the total cumulative distance `AL` and change in bearing `d.phi` as a broad attribute.  Finally total time for the trip, which excludes stops. All features are normalized so that they have a mean=0 and sd=10.

Using just these broad measures, let's compare two trips I manually matched earlier.  Here are the trips.

```{r ngram_compare, fig.width=10}
par.orig <- par(mfrow=c(1,2))

driver.ngrams <- data.frame()
for ( trip.id in c(44,45)) {
    trip <- getTrip(driver.id, trip.id, v.thresh=2) 
    plotTrip(trip)
    segs <- getSegments(trip)
    plotTripOverlay(trip, segs[segs$seg.type=="straight", ], col='yellow') #show straights
    plotTripOverlay(trip, segs[segs$seg.type=="turn", ], col='magenta') #show turns
    plotTrip.stopOverlay(trip, segs[segs$seg.type=="stop", ])
    
    driver.ngrams <- rbind( driver.ngrams, getNgrams( segs ))
}

```

Now lets plot scatter plots in 2D for the features we're looking at. 

The are several well matched segment pairs between the two trips (points from each trip that are close together) visible in the AL-d.phi plot.  This sets the stage for generalized matching, which I'll test on the following manually selected pairs from the scatter plot.
The first 9 are strong matches, and the next 4 have points that are close together on the AL-d.phi scatter plot, but do not look like similar segments.  I create a distance calculation that brings in the 3rd axis (time=tl for segment).  The index plot below shows a clear separation after pair #9 (segments 8/44 and 6/45)

Note: AL-d.phi  get's pretty close by itself, but has several false positives.  Finding the third axis (tl), took some trial and error.  Things I tried unsuccessfully were the ratio of AL/CL, modifying d.phi based on whether the 1st or last segmentm was short compared to AL (the exit angle from turns can be off by quite a bit, due to the lack of granuality of the data)

```{r ngram_scatter, fig.width=10}
pairs <- matrix( c (6,4   ,7,5   ,13,11   ,9,7   ,10,8  ,8,6   ,11,9   ,12,10   ,15,14  #match
                   ,5,13  ,4,13  ,16,9    ,14,9   #do not match
                   ), ncol=2, byrow=TRUE )
pairs <- cbind( pairs, rep(NA, nrow(pairs)))

plotPairEllipse <- function (scatter, feax, pair, kf=1.2, ...) {
    xyf1 <- with(scatter, as.numeric( scatter[ tripnum==44 & id==pair[1] , feax]))
    xyf2 <- with(scatter, as.numeric( scatter[ tripnum==45 & id==pair[2] , feax]))
    ellipse( xyf1[1], xyf1[2], xyf2[1], xyf2[2], kf=kf, new=FALSE, ... )
}

par(mfrow=c(1,2))
mar.orig <- par(mar=c(4,4,1,1))

scatter <- select(driver.ngrams, driver, tripnum, id, AL, CL, d.phi, tl)
for(i in 4:ncol(scatter)) scatter[,i] <- scaleFeatures(scatter[,i], colnames(scatter)[i])
scatter$col <- ifelse( scatter$tripnum == 45, "blue", "red")
scatter$pos <- ifelse( scatter$tripnum == 45, 1, 3)
scatter$pch <- ifelse( scatter$tripnum == 45, 1, 2)

feax <- c("AL","d.phi")
with(scatter, plot(scatter[,feax], asp=1, col=col, pch=pch))
with(scatter, text(scatter[,feax], labels=id, col=col, pos=pos))
for (pid in 1:nrow(pairs)) plotPairEllipse( scatter, feax, as.integer(pairs[pid,(1:2)]), kf=2, lty=2 )

feax <- c("AL", "tl")
with(scatter, plot(scatter[,feax], asp=1, col=col, pch=pch))
with(scatter, text(scatter[,feax], labels=id, col=col, pos=pos))
for (pid in 1:nrow(pairs)) plotPairEllipse( scatter, feax, as.integer(pairs[pid,(1:2)]), kf=2, lty=2 )


for (i in 1:nrow(pairs) )  {
    pairs[i,3] <- compareNgrams( 2591, 44, pairs[i,1], 45, pairs[i,2], show.plot=FALSE, sf= scaleFeatures2  ) 
}
pairs <- pairs[ order( pairs[ ,3]), ]

plot(pairs[ ,3], type="b", ylab="distance")
plabels <- rep("", nrow(pairs))
plabels[9] <- paste( pairs[9, 1:2], collapse=",")
plabels[10] <- paste( pairs[10, 1:2], collapse=",")
text( pairs[ ,3], plabels, pos=3)

par(mfrow=c(2,2))
dist <- compareNgrams( 2591, 44, pairs[9 ,1], 45, pairs[9 ,2], show.plot=TRUE, sf= scaleFeatures2  )
dist <- compareNgrams( 2591, 44, pairs[10,1], 45, pairs[10,2], show.plot=TRUE, sf= scaleFeatures2  )

```

So to automate the approach, I take a look at all the possible ngram combinations for this trip pair and calculate the `distance` in order to rank them.  The first plot below shows all possible cominations, however we're only interested in the best match for each segment, so the second plot shows a pruned version of the list which plots the difference only for the best matches of each distance.  There is a clear elbow in plot separating good matches from poor ones.

```{r ml_44_45, fig.width=10}
par(mfrow=c(1,2))
trip1.id <- 44
trip2.id <- 45

n.ng1 <- length( which( driver.ngrams$tripnum == trip1.id ))  #rows
n.ng2 <- length( which( driver.ngrams$tripnum == trip2.id ))  #cols

pairs <- data.frame( seg1=integer(), seg2=integer(), distance=numeric() )
for (i in 1:n.ng1 ) { 
    for(j in 1:n.ng2) {
        pair <- data.frame(  i, j, 
                distance = compareNgrams( 2591, trip1.id, i, trip2.id, j, show.plot=FALSE, sf= scaleFeatures2  )
                )
        pairs <- rbind( pairs, pair)
    }
}
pairs <- pairs[ order( pairs[ ,3]), ]

# now weed out anything that is doubly matched (first match stays)
pairs.pruned <- pairs[1, ]
for (i in 2:nrow(pairs)) {
    if ( pairs[i,1] %in% pairs.pruned[ ,1] 
       | pairs[i,2] %in% pairs.pruned[ ,2] ) {
        #cat ("remove row", i, as.integer(pairs[i,1:2]), "...\n")
    } else {
        pairs.pruned <- rbind( pairs.pruned, pairs[i, ])
    }
}

plot(pairs[ ,3], type="b", ylab="distance", main="All possible combinations")
plot(pairs.pruned[ ,3], type="b", ylab="distance", main="Best Matches")

plotPair <- function (pairs, rank) {
    pid <- rownames(pairs)[rank]
    compareNgrams( driver.id, trip1.id, pairs[pid,1], trip2.id, pairs[pid,2], sf=scaleFeatures2 , show.plot=TRUE)
}
```

Here are plots for the 9th and 10th ranked pairs.

```{r rank9_10, fig.width=10}
par(mfrow=c(2,2))
plotPair( pairs.pruned, 9 )
plotPair( pairs.pruned, 10 )

```

Notes: pair #10 should also be a match, but the straights weren't detected well.  Increasing the zone settings for parse.bearings (to 2.5) would probably improve things, but since it might break things elsewhere, I'm leaving it be.  The match detection is pretty good as is for this trip at least.   

```{r ngram10, fig.width=10}
par(mfrow=c(2,2))

quickPlot(2591, 44)
quickPlot(2591, 44, tx=90, tl=20)
quickPlot(2591, 45)
quickPlot(2591, 45, tx=40, tl=20)
```

The next best match is clearly not a match, so we're pretty close to the optimum.

```{r ngram11, fig.width=10}
par(mfrow=c(1,2))
plotPair( pairs.pruned, 11 )
```

For now, let's use `dist < 5` as the arbitrary cutoff for a good match.  Then 
```{r score1}
match.thresh <- 5
good.matches <- sum( pairs.pruned$distance < match.thresh )
good.matches

score.44 <- good.matches/n.ng1
score.45 <- good.matches/n.ng2
```
The scoring for the matches are  
* score 44 vs 45 = `r score.44`  
* score 45 vs 44 = `r score.45`  

where 0 is no matches and 1 is all matches.

## New functions used in these notes
```{r functions}
getSegments
getNgrams
quickPlot

```

